Canopy leaf area of an evergreen Eucalyptus woodland does not respond to elevated

### 2015-05-01 manuscript.Rmd
fixed 4 missing refs

\
This lack of a response to elevated $C_a$ counters our expectations and those in previous publications [@Norby2003; @Mccarthy2007], for a number of reasons. First, in a review of forest FACE sites, Norby and Zak [-@Norby2011] showed that low *L* stands were more responsive to $eC_a$. They argued that high *L* canopies would not benefit from further increases in *L*, because it would lead to only very small increases in total absorbed radiation. The mean *L* at the EucFACE was ca. `r mean(ba$LAI.mean)` $m^2$ $m^{-2}$, which is low compared to a global compendium of forest *L* [@Iio2013], and the lowest *L* of any forest FACE experiment [@Norby2011], but in line with a range of values for native *Eucalyptus* forests and woodlands reported by Whitehead and Beadle [-@Whitehead2004] and Iio et al. [-@Iio2013]. Based on this *L* estimate and the relationship across FACE sites reported by Norby and Zak [-@Norby2011], we could expect about a 30% increase in *L* under $eC_a$. It is clearly too early to draw general conclusions across FACE sites in terms of *L* response, moreover because water and nutrient likely play a central role in productivity and *L* dynamics at the EucFACE. Indeed, experimental evidence clearly demonstrates strong limitations of *L* by nutrients and water in Eucalyptus stands [@White2010].

@@

## Consequences for phenology models
Current-generation phenology models are heavily biased towards temperate conditions [@Primack2011], where leaf flushing occurs in spring in response to increase in temperature [@Richardsonet2012; @Savoy2015]. In contrast, phenology of evergreen *Eucalyptus* is more opportunistic, and shows rapid response to water availability (Fig. 7). Specht and Brouwer [-@Specht1975] demonstrated for four *Eucalyptus* species that leaf growth  peaked both in spring and autumn, with low winter temperatures and high summer temperatures (and low water availability) suppressing leaf growth. Similarly, Pook  [-@Pook1984a] demonstrated leaf production during most months in *Corymbia maculata*, except during winter and suppression by dry periods. Thus a minimum temperature is also required for leaf growth to occur. We also have not observed leaf flushing in winter time, but at our site this was confounded with water availability as both winters during the study period were relatively dry (Fig. 7).

\
Although the response of *L* to water availability is qualitatively well-known, current-generation phenology models fall short in incorporating this response. In a comparison of six leading ecosystem and global vegetation models,  Powell et al. [-@Powell2013] showed that the steady decrease in *L* in an Amazon drought experiment could be matched by only one model. Simulations from De Kauwe et al. [-@DeKauwe2014] at ambient and elevated $C_a$ also showed little response to inter-annual variation in rainfall (pers. comm. Martin De Kauwe). In most models  that provide prognostic estimates of *L* [i.e., excluding models that take *L* as an input parameter; e.g. @Kala2013), temporal changes in *L* are modelled by first calculating total net photosynthetic productivity (carbon availability, gross photosynthesis minus respiration), and allocating some fraction to the leaf pool. Combined with - usually constant - leaf turnover and specific leaf area (ratio of leaf area to mass), this yields predictions of *L* in response to environmental drivers such as $eC_a$ (Fig. 2). We hypothesize that this algorithm is not well suited to explain the dynamic nature of *L* in response to water availability observed in evergreen broad-leaved forests, because not all important mechanisms are incorporated in these models. In order to construct a phenology model where phenology is responsive to water availability, it should include responses of both leaf turnover and leaf growth to fluctuating water availability. Water limitation is known to directly influence leaf turnover [@Pook1997], possibly to maintain sufficient rates of water availability to the remaining leaves [@Vilagrosa2003; @Misson2010], and rates of leaf expansion [@Muller2011]. Some existing ecosystem models allow for increased leaf turnover during periods of low water availability (e.g. G’DAY, see Corbeels et al. 2005), but this process has been poorly quantified.

>>>>

\
This lack of a response to elevated $C_a$ counters our expectations and those in previous publications [@Norby2003; @McCarthy2007], for a number of reasons. First, in a review of forest FACE sites, Norby and Zak [-@Norby2011] showed that low *L* stands were more responsive to $eC_a$. They argued that high *L* canopies would not benefit from further increases in *L*, because it would lead to only very small increases in total absorbed radiation. The mean *L* at the EucFACE was ca. `r mean(ba$LAI.mean)` $m^2$ $m^{-2}$, which is low compared to a global compendium of forest *L* [@Iio2013], and the lowest *L* of any forest FACE experiment [@Norby2011], but in line with a range of values for native *Eucalyptus* forests and woodlands reported by Whitehead and Beadle [-@Whitehead2004] and Iio et al. [-@Iio2013]. Based on this *L* estimate and the relationship across FACE sites reported by Norby and Zak [-@Norby2011], we could expect about a 30% increase in *L* under $eC_a$. It is clearly too early to draw general conclusions across FACE sites in terms of *L* response, moreover because water and nutrient likely play a central role in productivity and *L* dynamics at the EucFACE. Indeed, experimental evidence clearly demonstrates strong limitations of *L* by nutrients and water in Eucalyptus stands [@White2010].

@@

## Consequences for phenology models
Current-generation phenology models are heavily biased towards temperate conditions [@Primack2011], where leaf flushing occurs in spring in response to increase in temperature [@Richardson2012; @Savoy2015]. In contrast, phenology of evergreen *Eucalyptus* is more opportunistic, and shows rapid response to water availability (Fig. 7). Specht and Brouwer [-@Specht1975] demonstrated for four *Eucalyptus* species that leaf growth  peaked both in spring and autumn, with low winter temperatures and high summer temperatures (and low water availability) suppressing leaf growth. Similarly, Pook  [-@Pook1984a] demonstrated leaf production during most months in *Corymbia maculata*, except during winter and suppression by dry periods. Thus a minimum temperature is also required for leaf growth to occur. We also have not observed leaf flushing in winter time, but at our site this was confounded with water availability as both winters during the study period were relatively dry (Fig. 7).

\
Although the response of *L* to water availability is qualitatively well-known, current-generation phenology models fall short in incorporating this response. In a comparison of six leading ecosystem and global vegetation models,  Powell et al. [-@Powell2013] showed that the steady decrease in *L* in an Amazon drought experiment could be matched by only one model. Simulations from De Kauwe et al. [-@DeKauwe2014] at ambient and elevated $C_a$ also showed little response to inter-annual variation in rainfall (pers. comm. Martin De Kauwe). In most models  that provide prognostic estimates of *L* [i.e., excluding models that take *L* as an input parameter; e.g. @Kala2013), temporal changes in *L* are modelled by first calculating total net photosynthetic productivity (carbon availability, gross photosynthesis minus respiration), and allocating some fraction to the leaf pool. Combined with - usually constant - leaf turnover and specific leaf area (ratio of leaf area to mass), this yields predictions of *L* in response to environmental drivers such as $eC_a$ (Fig. 2). We hypothesize that this algorithm is not well suited to explain the dynamic nature of *L* in response to water availability observed in evergreen broad-leaved forests, because not all important mechanisms are incorporated in these models. In order to construct a phenology model where phenology is responsive to water availability, it should include responses of both leaf turnover and leaf growth to fluctuating water availability. Water limitation is known to directly influence leaf turnover [@Pook1997], possibly to maintain sufficient rates of water availability to the remaining leaves [@Vilagrosa2003; @Misson2011], and rates of leaf expansion [@Muller2011]. Some existing ecosystem models allow for increased leaf turnover during periods of low water availability (e.g. G’DAY, see Corbeels et al. 2005), but this process has been poorly quantified.
